Anxiety: a decision-theoretic perspective

### 2019-01-31 manuscript/manuscript.tex
sections 4,5

Basically reviewing the arguments of Huys here \citep{Huys2012, Lally2017}.
Need to get facts straight before going through all of this.

@@

>>>>

In this next section, we review the literature on aversive pruning. In large,
multi-step sequential decision environments, the space of all possible sequences
of choices grows exponentially larger with increasing sequence length. As such,
it is infeasible for decision making agents to exhaustively explore all possible
sequences select that which maximizes reward from the entire set. Heuristics for
narrowing the search space then are not only plausible but beneficial for a
decision making agent.

One such heuristic that has been proposed in the literature is aversive pruning
\citep{Huys2012}. Aversive pruning is defined as a Pavlovian response to encountering
a large loss in planning such that sequences involving large losses are discarded
from further evaluation. An example of scenario is presented in Figure 4. The prediction
in such an environment is that an agent would discard the branch of the decision
tree involving an immediate large negative loss even though this branch objectively
contains the reward maximizing (loss minimizing) sequence of choices.

\cite{Huys2012} (and later \cite{Lally2017}) tested this prediction in the
choice envrionment displayed in Figure 4. Participants received extensive training
in this decision tree environment (first without outcomes and then later with
rewards) in order to facilitate planning. In free-planning trials, many participants
exhibited this Pavlovian bias, rejecting choice sequences which involved the large
loss even when the sequence was objectively reward maximizing (loss minimizing).
Importantly \cite{Huys2012} and \cite{Lally2017} found the degree of aversive
pruning was correlated with depressive and anxiety symptoms, respectively.

The aversive pruning principle makes tremendous sense and draws on a robust literature
of heuristics and computational shortcuts for circumventing computationally intractable
cognitive processes. It is important to note, however, that a control interpretation
also makes similar predictions. Figure 4b-d shows the preferred routes of simulated
agents in the same environments under increasingly pessimistic expectations of
future actions. As can be observed, pessimistic agents are similarly likely to avoid
the branch with the large loss, though for different reasons. Under the pessimistic case,
the agent is increasingly less confident that the large gains later in
sequence will be realized; as such, the agent is less confident that the initial
large loss will be offset. Consequently, pessimistic agents would prefer the branch
with smaller losses (even if this is objectively suboptimal). This prediction
is similarly consistent with the finding that avoidance of the initial large loss
is correlated with anxiety \cite{Lally2017}.

An interpretation like this would be consistent with reduced beliefs in one's
own competency in such an environment. This is not strictly surprising. Planning
multi-choice sequences even in relatively simple environments like those used
in these experiments still ostensibly require working memory ability to keep in
mind the possible consequences along each branch. Insofar that working memory
ability is disrupted in anxiety disorders \citep{Moran2016}, a pessimistic belief
may be justified and possibly optimal.

It is also worht noting that the two theories can be easily disentangled with a
simple manipulation of the choice environment. The two theories predict
different choices when the large loss comes later in sequence. For aversive pruning,
late large losses should have little effect on choice behavior. For pessimistic
learning, however, the value of large losses should back-propagate and contaminate
its associated branch. Future experiments can tease these apart as tests of the theory.

@@

If an agent is confident that their compentency in decision-making in intact,
then it follows that situations in which an individual is able to make a choice
should be preferable to situations in which an individual is not. At worst,
an agent is no better off than if they had not made a choice; at best, an agent
is better able to steer themselves toward desirable outcomes. Such is the logic
that underlies an increasingly robust literature suggesting choice is inherently
valuable \citep{Leotti2010}.

This free choice premium has been demonstrated multiple times in humans \citep{Suzuki1997,Leotti2011,Leotti2014,Cockburn2014} using a variety of decision
paradigms. For the present purposes, we will describe the experiments presented in
\citep{Leotti2011,Leotti2014}. In these experiments, participants complete a
two-stage decision making task (Figure 5). In the first stage, participants are
make a choice between two cues: one leading to free choice and the other leading
to a fixed choice. In the second stage, participants are able to choose between
a second set of cues (free choice) or are forced to choose a cue (fixed choice).
Unbeknownst to participants, all cues yield reward following the same outcome
distribution; thus, no cue is advantageous in the long-run. The free choice
premium is measured as a preference for the cue leading to free choice. \cite{Leotti2011,Leotti2014}
find participants still prefer the free choice cue, despite this conferring no
objective benefit to participants.

Insofar that this effect is contingent on the belief of one's own competency,
it stands to reason that this free choice premium should be absent in individuals
with anxiety and anxious control beliefs. We demonstrate this effect in Figure 5.
In each trial, the reward outcomes are equal for all options (-1, 1). In individuals
with increasingly pessimistic beliefs, the worst outcome in free choice is increasingly
back-propagated to the free-choice cue, thereby making it less attractive than
the fixed choice cue. Hence, pessimistic learners, on average, do not exhibit
this choice bias.

To our knowledge this prediction has not been empirically tested.
